---
title: "Class 09"
author: "Kate Ruiz (PID A17671200)"
format: pdf
---

# 2. Importing Candy Data



```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```

## 2.1 What is in this data set?

> Q1. How many different candy types are in this dataset?

There are 85 different candies.


```{r}
dim(candy)
```

```{r}
nrow(candy)
```
```{r}
sum(candy$fruity)
```


> Q2. How many fruity candy types are in the dataset?

There are 38 fruity candy types in this data set.

## 2.2 What is your favorite candy? 

```{r}
candy["Twix", ]$winpercent
```

```{r}
library(dplyr)

candy |> 
  filter(row.names(candy)=="Twix") |> 
  select(winpercent)
```

> Q3.  What is your favorite candy (other than Twix) in the dataset and what is it’s winpercent value?

The candy I chose other than Twix is Almond Joy.  The winpercent value is 50.3475.

```{r}
candy["Almond Joy", ]$winpercent
```

> Q4. What is the winpercent value for “Kit Kat”?

The winpercent value for Kit Kat is 76.7686.
 
```{r}
candy["Kit Kat", ]$winpercent
```
 
> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

The winpercent value for tootsie roll snack bars is 49.6535.

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

### side note: the skimr::skim() function

```{r}
library("skimr")
skim(candy)
```
> Q6.  Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

winpercent seems to be on a different scale.  It looks like a percentage as opposed to a proportion.

> Q7. What do you think a zero and one represent for the candy$chocolate column?

I think it's a binary for yes and no.  0 is likely no and 1 is likely yes.
```{r}
skim(candy$chocolate)
```

# 3 Exploratory Analysis

> Q8. Plot a histogram of winpercent values using both base R an ggplot2.

```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)
ggplot(candy, aes(x=winpercent)) +
  geom_histogram()
  
```

> Q9. Is the distribution of winpercent values symmetrical?

It does not look very symmetrical.  It looks skewed to the right.

> Q10. Is the center of the distribution above or below 50%?

```{r}
median(candy$winpercent)
```
The center of the distribution is below 50% at 47.83%.

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
fruity_candy <- candy$winpercent[as.logical(candy$fruity)]
```

```{r}
chocolate_candy <- candy$winpercent[as.logical(candy$chocolate)]
```

```{r}
mean(fruity_candy)
```
```{r}
mean(chocolate_candy)
```
On average, chocolate candy is more highly ranked (60.92%) than fruity candy (44.12%).  

> Q12. Is this difference statistically significant?

```{r}
t.test(chocolate_candy, fruity_candy)
```
The difference is statistically significant as the p value is less than or equal to .05 (p=2.871e-08).

# 4 Overall Candy Rankings

> Q13. What are the five least liked candy types in this set?

The five least liked candy types in this set are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters.

```{r}
candy |> arrange(winpercent) |> head(5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

The 5 all time favorite candies out of this data set are Snickers, Kit Kat, Twix, Reese's Miniatures, and Reese's Peanut Butter cup.

```{r}
candy |> arrange(winpercent) |> tail(5)
```


> Q15.Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col() + 
  theme(axis.text.y = element_text(size = 3))
```

## 4.1 Time to add some useful color

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) +
  theme(axis.text.y = element_text(size = 3))
```

> Q17. What is the worst ranked chocolate candy?

The worst ranked chocolate candy are sixlets.
> Q18. What is the best ranked fruity candy?

The best ranked fruity candy are starburst.

# 5 Taking a look at pricepercent

```{r}
library(ggrepel)

# How about a plot of win vs price
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Tootsie Roll Midgies  have the highest winpercent for the lowest pricepercent. 
```{r}
best <- candy$winpercent / candy$pricepercent

candy[which.max(best), c("winpercent", "pricepercent")]

```

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

The tip 5 most expensive candy types are Nik L Nip, Nestle Smarties, Ring Pop, Hershey's Krackel, and Hershey's Milk Chocolate.  Of these the least popular is Nik L Nip.

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

# 6 Exploring the correlation structure

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

It seems like fruity and a variety of things.  Of note are fruity and chocolate, fruity and bar, and pluribus and bar (among  others).
> Q23. Similarly, what two variables are most positively correlated?

It looks like it would be chocolate and bar.  It seems many chocolate + XX pairs are blue (positively correlated).


# 7 Principal Component Analysis

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)
```

```{r}
plot(pca$x[ ,1:2])
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```


```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```


```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)
ggplotly(p)
```

```{r}
ggplot(pca$rotation) +
  aes(x = PC1, y = reorder(rownames(pca$rotation), PC1)) +
  geom_col()
```

> Q24. Complete the code to generate the loadings plot above. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you? Where did you see this relationship highlighted previously?

In the positive direction, fruity, pluribus, and hard are picked up.  It feels a little confusing.  They were negatively/ anticorrelated previously.

# 8 Summary

> Q25. Based on your exploratory analysis, correlation findings, and PCA results, what combination of characteristics appears to make a “winning” candy? How do these different analyses (visualization, correlation, PCA) support or complement each other in reaching this conclusion?

The winning candy has chocolate, nougat, bars, and peanutyalmondy. The visualizations allowed us to look at costs and like-ability.  The correlations helped us see the best combinations that led to better liking.  PCA provided predictive powers.