---
title: "class12"
author: "Kate Ruiz (PID: A17671200)"
format: html
---

# Section 1: Proportion of G/G in a population

Downloaded a CSV file from Ensemble < https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=959672880#373531_tablePanel

Here we read this CSV file

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```


```{r}
(table(mxl$Genotype..forward.strand.)/nrow(mxl)) * 100
```

Now let's look at a different population.  I chose what the professor chose, which is GBR.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(gbr)
```

Find portion of GIG 

```{r}
round(table(gbr$Genotype..forward.strand.)/nrow(gbr)*100,2)
```

This variant that is associated with childhood asthma is more frequent in the GBR population than the MKL population.  

Let's now dig into this further.

## Section 4: Population Scale Analysis
One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. 
So, you processed about ~230 samples and did the normalization on a genome level. Now, you want to find whether there is any association of the 4 asthma-associated SNPs (rs8067378...) on ORMDL3 expression. 

How many samples do we have?

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```





