---
title: "Class014"
author: "Kate Ruiz (PID: A17671200)"
format: pdf
---

## Background

The data for today's mini project comes from a knock-down study of an important HOX gene.

## Data Import

```{r}
library(DESeq2)
```

```{r}
countData <- read.csv("GSE37704_featurecounts.csv", row.names=1)
colData <- read.csv("GSE37704_metadata.csv", row.names=1)
```

```{r}
head(colData)
```

```{r}
head(countData)
```

Have a wee peek at these:

### Clean Up (data tidyig)

We need to remove the funny "length" column from out `countData` to make the columns match the rows in `colData`.

> Q Complete the code below to remove the troublesome first column from countData

> Q Complete the code below to filter countData to exclude genes (i.e. rows) where we have 0 read count across all samples (i.e. columns).

Tip: What will rowSums() of countData return and how could you use it in this context?
```{r}
countData <- countData[,-1]
```

Check macth of `colData` and `countData`

```{r}
rownames(colData) == colnames(countData)
```

## DESeq Analysis

```{r}
library(DESeq2)
```

```{r}
head(countData)
```

### Remove zero count genes

```{r}
to.keep <- rowSums(countData) > 0
countData <- countData[to.keep,]

```

```{r}
library(DESeq2)
```

```{r}
dds<- DESeqDataSetFromMatrix(countData=countData,
                       colData=colData,
                       design=~condition)
```

```{r}
dds = DESeq(dds)
dds
```

```{r}
res = results(dds)

```

> Q  Call the summary() function on your results to get a sense of how many genes are up or down-regulated at the default 0.1 p-value cutoff.

```{r}
summary(res)
```

### Setting up input object

### Running DESeq

### Getting results

## Volcano plot

A plot of log2 fold change vs -log of Adjusted P value

```{r}
library(ggplot2)

ggplot(res)+
  aes(log2FoldChange,
      -log(padj))+
  geom_point()
```

## Add annotation
```{r}
ggplot(res)+
  aes(log2FoldChange,
      -log(padj))+
  geom_point()+
  geom_hline(yintercept=-log(0.01), col="red")
```

```{r}
mycols <- rep("gray", nrow(res))
mycols[abs(res$log2FoldChange)>2] <- "darkgreen"
mycols[res$padj >= 0.01] <- "gray"

ggplot(res)+
  aes(log2FoldChange,
      -log(padj))+
  geom_point(col=mycols)+
  geom_hline(yintercept=-log(0.01), col="red")

```


```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

columns(org.Hs.eg.db)

res$symbol = mapIds(org.Hs.eg.db,
                    keys=rownames(res), 
                    keytype="ENSEMBL",
                    column="SYMBOL")

res$entrez = mapIds(org.Hs.eg.db,
                    keys=rownames(res),
                    keytype="ENSEMBL",
                    column="ENTREZID")



head(res)
```

```{r}
library(gage)
library(gageData)
library(pathview)
```
We need a foldchanges vector for `gage()`

```{r}
foldchanges <- res$log2FoldChange
names(foldchanges)<-res$entrez
```

## Pathway Analysis

### KEGG
```{r}
library(pathview)
library(gage)
library(gageData)

data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
```

```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```
```{r}
# Get the results
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

```{r}
attributes(keggres)
```

```{r}
# Look at the first few down (less) pathways
head(keggres$less)

```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```

```{r}
# A different PDF based output of the same data
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```
```{r}
## Focus on top 5 upregulated pathways here for demo purposes only
keggrespathways <- rownames(keggres$greater)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids

```
```{r}
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")
```







### GO
```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets)

lapply(gobpres, head)

```




### Reactome
```{r}
sig_genes <- res[res$padj <= 0.05 & !is.na(res$padj), "symbol"]
print(paste("Total number of significant genes:", length(sig_genes)))
```

```{r}
write.table(sig_genes, file="significant_genes.txt", row.names=FALSE, col.names=FALSE, quote=FALSE)

```


















