---
title: "Class17"
author: "Kate Ruiz (PID: A17671200)"
format: pdf
---

## Core Unix commands


Most unix commands have super short names, which makes them quick to type 
but annoying to learn.  Major file system related commands include:
pwd: Print working directory
ls: list files and directories
cd: change directory
mkdir: make a new directory
rm: remove diles and directories (delete) super dangerous
cp: copy files (source > destination)
mv: move files or directories (basically rename)
nano: a wee text command editor (very basic but always available)

## AW EC2 Instance
Connect to my instance with:

ssh -i ~/Downloads/bimm143_kate.pem ubuntu@ec2-34-219-54-212.us-west-2.compute.amazonaws.com

Secure Copy files between machines, in this case from our instance to our laptop

scp -i ~/Downloads/bimm143_kate.pem ubuntu@ec2-34-219-54-212.us-west-2.compute.amazonaws.com:/home/ubuntu/work/bimm143_github/class16/results.txt .

## PCA 

```{r}
library(tximport)

# setup the folder and file-names to read
folders <- dir(pattern="SRR21568*")
samples <- sub("_quant", "", folders)
files <- file.path( folders, "abundance.h5" )
names(files) <- samples

txi.kallisto <- tximport(files, type = "kallisto", txOut = TRUE)
```


```{r}
head(txi.kallisto$counts)
```

```{r}
getwd()

```

```{r}
list.files()
```

```{r}
folders <- dir(pattern = "SRR21568")
folders
```


```{r}
samples <- sub("_quant", "", folders)
samples
```


```{r}
files <- file.path(folders, "abundance.h5")
files
```

```{r}
names(files) <- samples
files
```

```{r}
txi.kallisto <- tximport(files, type = "kallisto", txOut = TRUE)
```

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("rhdf5")
```

```{r}
colSums(txi.kallisto$counts)
```

```{r}
sum(rowSums(txi.kallisto$counts)>0)
```


```{r}
to.keep <- rowSums(txi.kallisto$counts) > 0
kset.nonzero <- txi.kallisto$counts[to.keep,]
```


```{r}
keep2 <- apply(kset.nonzero,1,sd)>0
x <- kset.nonzero[keep2,]
```


```{r}
pca <- prcomp(t(x), scale=TRUE)
```

```{r}
summary(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2],
     col=c("blue","blue","red","red"),
     xlab="PC1", ylab="PC2", pch=16)
```

```{r}
names(pca)
```

```{r}
head(pca$x)
```

```{r}
df <- as.data.frame(pca$x)
head(df)
```

```{r}
rownames(pca$x)
```

```{r}
colnames(df)
head(df)
```

```{r}
df$condition <- c("A", "A", "B", "B")
head(df)
```

```{r}
head(df)
```

```{r}
library(ggplot2)
ggplot(df, aes(x = PC1 , y = PC2 , color = condition)) +
  geom_point()
```

```{r}
ggplot(df, aes(x = PC1 , y = PC3 , color = condition)) +
  geom_point()
```

```{r}
ggplot(df, aes(x = PC2 , y = PC3 , color = condition)) +
  geom_point()
```

