---
title: "class06: R projects"
author: "Kate Ruiz (PID: A17671200)"
format: pdf
toc: true
---

## Background

Functions are at the heart of using R.  Everything we do involves calling and using functions (from data input, analysis to results output).  


All functions in R have at least 3 things: 

1. A **name** the thing we use to call the function.
2. One or more input  **arguments** that are comma separated 
3. The **body**, lines of code between curly brackets {} that does the work of the function.

## A first function

Let's write a silly wee function to add some numbers:
```{r}
add <- function(x) {
  x + 1 
}
```

Let's try it out 
```{r}
add(100)
```

Will this work

```{r}
add(c(100,200,300))
```
Modify to be more useful and add more than just 1

```{r}
add <- function(x, y=1) {
  x + y 
}
```


```{r}
add(100, 10)
```
Will this work?
```{r}
add(100,0)
```

```{r}
plot(1:10, col="blue", typ="b")
```

```{r}
log(10, base=10)
```

> **Note to Kate** Input arguments can be either **required** or **optional**.  The later have a fall-back default that is specified in the function code with an equals sign.


```{r}
#add(x=100, y=200, z=300)

```

## A second function 

all functions in R look like this


```
name <- function(arg) {
  body
}
```
The `sample()` function in R ... 

```{r}
sample(1:10, size=4)
```

> Q. Return 12 numbers picked randomly from the input 1:10

```{r}
sample(1:10, size=12, replace=TRUE)
```

> Q. Write the code to generate a random 12 nucleotide long DNA sequence?

```{r}
nucleotide <- c("A", "T", "G", "C")
sample(nucleotide, size=12, replace=TRUE)
```

> Q.  Write a first version function called `generate_dna()` that generates a user specified length `n` random DNA sequence?


```
name <- function(arg) {
  body
}
```

```{r}
generate_dna <-function(n=6){
  nucleotide <- c("A", "T", "G", "C")
  sample(nucleotide, size=n, replace=TRUE)
}
```

```{r}
generate_dna(100)
```


> Q. Modify your function to return a FASTA like sequence so rather than [1] "G" "T" "T" "G" "T" "C" "G" "A" "G" "G" "A" "G" we want "GTTG..."

```{r}
generate_dna <-function(n=6){
  nucleotide <- c("A", "T", "G", "C")
  sally<-sample(nucleotide, size=n, replace=TRUE)
  sally <- paste(sally, collapse="")
  return(sally)
}
```
## by default the last thing you type is what you get

```{r}
generate_dna(10)
```

> Q *bingus is a hairless cat*  Give the user an option to return FASTA format output ssequence or standard multi-element vector format

```{r}

generate_dna <-function(n=6, fasta=TRUE){
  nucleotide <- c("A", "T", "G", "C")
  sally<-sample(nucleotide, size=n, replace=TRUE)
  
  if(fasta){
    sally <- paste(sally, collapse="")
    cat("Helllooooooo")

  }
  return(sally)
}
```

```{r}
generate_dna(10, fasta=T)

```

### A new col function

> Q. Write a function called `generate_protein()` that generates a user specified length protein sequence in FASTA like format?

> Q. Use your new `generate_protein()` function to generate sequences between length 6 and 12 amino acids in length and check if any of these are unique in nature (i.e. found in the NR database at NCBI)

```{r}
generate_protein <- function(n, fasta=TRUE){
  amino_acids <- c("A","R","N","D","C",
  "E","Q","G","H","I",
  "L","K","M","F","P",
  "S","T","W","Y","V")
  more<-sample(amino_acids, size=n, replace=TRUE)
  if(fasta){
    more<-paste(more, collapse ="")
  }
  return(more)
}
```


```{r}
generate_protein(6, T)
generate_protein(7, T)
generate_protein(8, T)
generate_protein(9, T)
generate_protein(10, T)
generate_protein(11, T)
generate_protein(12, T)

```

Or we could do a `for()` loop:

```{r}
for (i in 6:12) {
  cat(i, "\n")
}
```

```{r}
for (i in 6:12) {
  cat(">", i,sep="", "\n")
  cat(generate_protein(i), "\n")
}
```

>id
AGKRT



