---
title: "Class 7: Machine Learning 1"
author: "Kate Ruiz (PID: A17671200)"
format: pdf
---

## Background
Today we will begin our exploration of some important machine learning methods, namely **clustering** and **dimensionality reduction**.

Let's make up some input data for clustering where we know what the natural "clusters" are.

The function `rnorm()` can be useful here.

```{r}
hist(rnorm(5000, mean=10))
```

> Q. Generate 30 random numbers at +3

```{r}
tmp <- c(rnorm(30, mean=3),
         rnorm(30, mean=-3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


## K-means clusetering

The main function in "base R"for K-means is called `kmeans()`

```{r}
km<-kmeans(x,centers=2)
km
```

>Q. What component of the results object details the cluster sizes?

```{r}
km$size
```
>Q. What component of the results object details the cluster centers?

```{r}
km$centers
```

>Q. What component of the results object details the cluster membershiip vector (i.e. our main result of which points lie in the cluster)?

```{r}
km$cluster
```

>Q. Plot our clustering results with points colored by cluster and alsp add the cluster centers as new colored blue?


```{r}
plot(x, col=km$cluster)
points(km$centers, col="blue", pch=15)
```

>Q. Run `kmeans()` again and this time produce 4 clusters (and call your result object `k4`) and make a results figure like above?

```{r}
k4 <- kmeans(x, centers=4)
plot(x, col=k4$cluster)
points(k4$centers, col="blue", pch=15)
```




The metric

```{r}
km$tot.withinss
k4$tot.withinss
```

>Q. Let's try different number of K (centers) from 1 to 30 and see what the best result is?

```{r}
ans <- NULL
for(i in 1:30) {
  ans <- c(ans, kmeans(x, centers=i)$tot.withins)}
ans
```

```{r}
plot(ans, type="o")
```
The problem is that it is self-fulfilling.  K-means will impose a clustering structure on your data even if it is not there-- it will always give you the answer you asked for even if that answer is silly!

## Hierarchical Clustering

The main function for Hierarchical Clustering is called `hclus()`
Unlike `kmeans()` (which does all the work for you), you can't just pass `hclust()` our raw input data.  It needs a "distance matrix" like the one returned from the `dist()` function.  

```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
```


To extract our cluster membership vector from a `hclust()` result object we have to "cut" our tree at a given height to yield separate "groups"/"branches".

```{r}
plot(hc)
abline(h=8, col="red", lty=2)
```

To di this we use the `cutree()` function on our `hclust()1 object:

```{r}
grps <- cutree(hc, h=8)
grps
```


```{r}
table(grps, km$cluster)
```


## PCA of UK food data

Import the dataset of food consumption in the UK:

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
x
```

> Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
dim(x)
```

```{r}
#rownames(x)
x[,1]
```

```{r}
rownames(x)<- x[,1]
```

To remove the first column I can use the minus index trick

```{r}
x <- x[,-1]
x
```



A better way to do this is to set the row names to the first column by with `read.csv()`

```{r}
x <- read.csv (url,row.names=1)
x
```


> Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

The second one.  I don't lose data.  I like fixing things at the source.

### Spotting major differences and trends

is difficult in this wee 17D dataset...

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```


```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```


### Pairs plots and heatmaps
```{r}
pairs(x, col=rainbow(nrow(x)), pch=16)
```

```{r}
library(pheatmap)
pheatmap( as.matrix(x) )
```

## PCA to the rescue

The main PCA function in "base R" is called `prcomp()`.  This function wants the transpose of our food data as input (i.e. the foods as columns and the countries as rows).

```{r}
pca <- prcomp(t(x))
```

```{r}
summary(pca)
```

```{r}
attributes(pca)
```

To make one of our main PCA result figures we turn to `pca$x` the scores along our new PCs.  This is called "PC plot" or "score plot" or "ordination plot" ...

```{r}
pca$x
```

```{r}
my_cols <- c("orange", "red", "blue", "darkgreen")
```



```{r}
library(ggplot2)

ggplot(pca$x)+
  aes(PC1, PC2)+
  geom_point(col=my_cols)
```


The second major result figure is called a "loadings plot" of "variable contributions plot" or "weight plot"

```{r}
ggplot(pca$rotation)+
  aes(PC1, rownames(pca$rotation))+
  geom_col()
```





